---
title: "scRNA-seq analysis"
author: "Toshiyuki Itai"
date: "6/10/2022"
output:
  html_document: default
  pdf_document: default
---
```{r}
#Clean the environment (if need be)
rm(list=ls())
```

```{r}
#Move to the analysis directory
setwd("/Users/titai/Documents/Research_Projects/Cleft/0_single_cell_analysis")
dir()
```

```{r}
#Load necessary packages
library(dplyr)
library(Seurat)
library(patchwork)
```


#Setup the Seurat Object
Load the dataset
```{r}
#FB00001024 (C57LJ) sc-RNA seq data
e11.5.data <- Read10X("/Users/titai/Documents/Research_Projects/Cleft/0_single_cell_analysis/E11.5/")
```

Initialize the Seurat object with the raw (non-normalized data).
I set the parameters basically referring to the serurat tutorial (https://satijalab.org/seurat/articles/pbmc3k_tutorial.html)
```{r}
e11.5 <- CreateSeuratObject(counts = e11.5.data, project = "e11.5", min.cells = 3, min.features = 200) 
e11.5
```

#Standard pre-processing workflow
## QC and selecting cells for further analysis
```{r}
# The [[ operator can add columns to object metadata. This is a great place to stash QC stats
e11.5[["percent.mt"]] <- PercentageFeatureSet(e11.5, pattern = "^mt-")
```

```{r}
# Visualize QC metrics as a violin plot
VlnPlot(e11.5, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```
```{r}
# FeatureScatter is typically used to visualize feature-feature relationships, but can be used
# for anything calculated by the object, i.e. columns in object metadata, PC scores etc.

plot1 <- FeatureScatter(e11.5, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(e11.5, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1 + plot2
```

To exclude potential doublets, apoptotic or lysed cells, cells with a nUMI larger than 80,000 or a percentage of UMI representing mitochondrial-encoded genes (pct.mito) of more than 10% were filtered out. 
we retained 7893 cells for further analysis.
```{r}
e11.5 <- subset(e11.5, subset = nCount_RNA < 80000 & percent.mt < 10)
e11.5
```

## Normalizing the data
```{r}
e11.5 <- NormalizeData(e11.5)
```

## Identification of highly variable features (feature selection)
```{r}
e11.5 <- FindVariableFeatures(e11.5, selection.method = "vst", nfeatures = 2000)
```

```{r}
# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(e11.5), 10)
```

```{r}
# plot variable features with and without labels
plot1 <- VariableFeaturePlot(e11.5)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot2
```

## Scaling the data
```{r}
all.genes <- rownames(e11.5)
e11.5 <- ScaleData(e11.5, features = all.genes)
```

Assign Cell-Cycle Scores
```{r}
#I downloaded the human cell-cycle scores from CellCycleScoreng() in Seurat and converted the into mouse gene symbols
s.genes <- read.csv("s.genes_mouse.csv")
g2m.genes <- read.csv("g2m_mouse.csv")
```

```{r}
s.genes <- as.vector(t(s.genes))
g2m.genes <- as.vector(t(g2m.genes))
```

```{r}
e11.5 <- CellCycleScoring(e11.5, s.features = s.genes, g2m.features = g2m.genes, set.ident = TRUE)
```

##Perform linear dimensional reduction
```{r}
e11.5 <- RunPCA(e11.5, features = VariableFeatures(object = e11.5))
```

```{r}
e11.5 <- RunPCA(e11.5, features = c(s.genes, g2m.genes))
```
```{r}
DimPlot(e11.5, reduction = "pca", group.by = "Phase")
#From the result below, I do regressing out of cell-cycele scores
```
## Regress out cell cycle scores during data scaling
```{r}
e11.5 <- ScaleData(e11.5, vars.to.regress = c("S.Score", "G2M.Score"), features = rownames(e11.5))
```

```{r}
#Check whether cell-cycle scores affects the characteristics of cells
e11.5 <- RunPCA(e11.5, features = c(s.genes, g2m.genes))
```
```{r}
#Plotting was not affected by cell-cygle signaling
DimPlot(e11.5, reduction = "pca", group.by = "Phase")
```
```{r}
#Run PCA again with high-variable features
e11.5 <- RunPCA(e11.5, features = VariableFeatures(object = e11.5))
```

```{r}
DimHeatmap(e11.5, dims = 1:20, cells = 500, balanced = TRUE)
```
## Determine the ‘dimensionality’ of the dataset
```{r}
# NOTE: This process can take a long time for big datasets, comment out for expediency. More
# approximate techniques such as those implemented in ElbowPlot() can be used to reduce
# computation time
e11.5 <- JackStraw(e11.5, num.replicate = 100)
e11.5 <- ScoreJackStraw(e11.5, dims = 1:20)
```

```{r}
JackStrawPlot(e11.5, dims = 1:20)
```

```{r}
ElbowPlot(e11.5)
```

```{r}
#Referring the results of heat map, JackStrawPlot, and ElbowPlot, I set the parmeter dims to 1:20.
e11.5 <- FindNeighbors(e11.5, reduction = "pca", dims = 1:20)
e11.5 <- FindClusters(e11.5, resolution = 0.5) #this parameter is the default parameters used in Seurat tutorial
```

```{r}
e11.5 <- RunUMAP(e11.5, dims = 1:20)
```

```{r}
a <- DimPlot(e11.5, reduction = "umap")
b <- DimPlot(e11.5, reduction = "umap", group.by = "Phase")

a+b
```

```{r}
#Annotation using marker genes
VlnPlot(e11.5, features = c("Alx1",     # mesenchyme
                            "Epcam",    # ectoderm
                            "Hemgn",    # redblood cells
                            "Cdh5",     # endothelium
                            "Fcer1g"))  # other blood cells
```
```{r}
# The results were similar to the original article
FeaturePlot(e11.5, features = c("Alx1", "Epcam", "Hemgn", "Cdh5", "Fcer1g"))
```

```{r}
#Cell annotation reffering to the above result
new.cluster.ids <- c("mesenchyme", #0
                     "mesenchyme", #1
                     "mesenchyme", #2
                     "mesenchyme", #3
                     "mesenchyme", #4
                     "ectoderm", #5 
                     "mesenchyme", #6
                     "ectoderm", #7 
                     "mesenchyme", #8
                     "mesenchyme", #9                    
                     "red blood cells", #10
                     "endothelium", #11
                     "ectoderm", #12
                     "ectoderm", #13
                     "other blood cells", #14
                     "mesenchyme", #15
                     "mesenchyme") #16
```

```{r}
names(new.cluster.ids) <- levels(e11.5)
e11.5 <- RenameIdents(e11.5, new.cluster.ids)
DimPlot(e11.5, reduction = "umap")
```

# Subclustering of ectoderm cells
```{r}
ect <- subset(e11.5, idents = c("ectoderm"))
```
```{r}
ect  #Similar result to the original paper (943 ectodermal cells)
```

```{r}
ect <- FindVariableFeatures(ect, selection.method = "vst", nfeatures = 2000)
```

```{r}
ect <- ScaleData(ect)
```

```{r}
ect <- RunPCA(ect)
```
```{r}
#Check wheter cell-cycle regression is necessary or not
ect <- RunPCA(ect, features = c(s.genes, g2m.genes))
DimPlot(ect, reduction = "pca", group.by = "Phase")
```
```{r}
ect <- ScaleData(ect, vars.to.regress = c("S.Score", "G2M.Score"), features = rownames(ect))
```
```{r}
#Cell-cycle does not affect cell characteristics.
ect <- RunPCA(ect, features = c(s.genes, g2m.genes))
DimPlot(ect, reduction = "pca", group.by = "Phase")
```


```{r}
ect <- RunPCA(ect, features = VariableFeatures(object = ect))
```

```{r}
DimHeatmap(ect, dims = 1:20, cells = 500, balanced = TRUE)
```
```{r}
ect <- JackStraw(ect, num.replicate = 100)
ect <- ScoreJackStraw(ect, dims = 1:20)
```
```{r}
JackStrawPlot(ect, dims = 1:20)
```
```{r}
ElbowPlot(ect)
```
```{r}
#I set these parameteres reffering to the original article and above results
#->  For reclustering of the ectodermal cells, 21 principal components and resolution 2.0 were used.
ect <- FindNeighbors(ect, dims = 1:20)
ect <- FindClusters(ect, resolution = 1.7)
```

```{r}
ect <- RunUMAP(ect, dims = 1:20)
DimPlot(ect)
```

```{r}
# e11 OE.2
VlnPlot(ect, features = c("Rprm", "Pcdh19", "Ctxn3"))
```
```{r}
# e0 OE.1 
VlnPlot(ect, features = c("Ebf1", "Mgp", "Lhfp"))
```
```{r}
#e4 anterior medial Nap
VlnPlot(ect, features = c("Fgf17", "Mecom", "Clu"))
```
```{r}
# e7 palate
VlnPlot(ect, features = c("Barx1", "Gm12446", "Dmrt2"))
```
```{r}
# e3 dental 
VlnPlot(ect, features = c("Shh", "Fgf8", "Lmo2"))
```
```{r}
# e10 fusion zone
VlnPlot(ect, features = c("Col9a1", "Adamts9", "Itga4"))
```
```{r}
# e5 NL groove 
VlnPlot(ect, features = c("Gjb6", "Barx2", "Gap43"))
FeaturePlot(ect, features = c("Gjb6", "Barx2", "Gap43"))
```
```{r}
# surface.2 
VlnPlot(ect, features = c("Robo2", "Wnt3", "Pcp4l1"))
FeaturePlot(ect, features = c("Robo2", "Wnt3", "Pcp4l1"))
```
```{r}
# surface 1
VlnPlot(ect, features = c("Wnt9b", "Wnt3", "Lmo1"))
FeaturePlot(ect, features = c("Wnt9b", "Wnt3", "Lmo1"))
```
```{r}
# e6 ambiguous
VlnPlot(ect, features = c("Slc39a1", "Spry2", "Slc24a5"))
FeaturePlot(ect, features = c("Slc39a1", "Spry2", "Slc24a5"))
```
```{r}
#e9 periderm
VlnPlot(ect, features = c("Gabrp", "Rhov", "Lypd3"))
FeaturePlot(ect, features = c("Gabrp", "Rhov", "Lypd3"))
```
```{r}
#38 ect & mes
VlnPlot(ect, features = c("Runx1t1", "Tnfaip2", "Snai1"))
FeaturePlot(ect, features = c("Runx1t1", "Tnfaip2", "Snai1"))
```

```{r}
ect.new.cluster.ids <- c(
"NL groove",#1
"Surface",#2
"OE.1",#3
"OE.2",#4
"Dental",#5
"Ect & mes",#6
"Anterior medial Nap",#7
"Fusion zone",#8
"Palate",#9
"Surface",#10
"Periderm",#11
"Surface" #12
)
```

```{r}
names(ect.new.cluster.ids) <- levels(ect)

ect <- RenameIdents(ect, ect.new.cluster.ids)
DimPlot(ect, reduction = "umap")
```
```{r}
save(e11.5, e11.5.data, ect, file = "C57LJ_scRNA_20220711.RData")
```




